---
title: 'Homework 1 - Research in Health Economics'
author: "Ammarah Ahmed"
date: ""
output: pdf_document 
    
---
```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, readr, 
               scales, kableExtra)
```

```{r, include=FALSE}
#l[oad("Hw1_workspace.Rdata")
```

# Question 1
```{r, include=FALSE}
full.ma.data<- readRDS("data/output/full_ma_data.rds")
```
```{r, include=FALSE}
total_obs<- as.numeric(count(full.ma.data %>% ungroup()))
```

```{r}
paste0("There are ", total_obs , " observations in the dataset")
```

# Question 2 
There are 26 different plans in the dataset.

# Question 3

```{r, include=FALSE}
plan.type.year1<- full.ma.data %>% group_by(plan_type, year) %>% count() %>% 
  arrange(year, -n) %>% filter(plan_type!="NA")
```
```{r, include=FALSE}
pe.year1<- pivot_wider(plan.type.year1, names_from = "year", values_from = "n", names_prefix = "Count_")
```

```{r}
options(knitr.kable.NA = 0)
knitr::kable(pe.year1, 
             col.names=c("Plan Type","2007", "2008", "2009",  "2010","2011","2012","2013","2014","2015"),
             format.args = list(big.mark=","), caption = "Plan Count by Year")
```

# Question 4

```{r, include=FALSE}
full.ma.data2<- full.ma.data %>% 
  filter(snp == 'No' & eghp == 'No' & !(planid %in% 800:899))
```
```{r, include=FALSE}
plan.type.year2<- full.ma.data2 %>% group_by(plan_type, year) %>% count() %>% 
  arrange(year, -n) %>% filter(plan_type!="NA")
```
```{r, include=FALSE}
pe.year2<- pivot_wider(plan.type.year2, names_from = "year", values_from = "n", names_prefix = "Count_")
```

```{r}
options(knitr.kable.NA = 0)
knitr::kable(pe.year2, 
             col.names=c("Plan Type","2007", "2008", "2009",  "2010","2011","2012","2013","2014","2015"),
             format.args = list(big.mark=","), caption = "Plan Count by Year")
```

# Question 5

```{r, include=FALSE}
contract.service.area <- readRDS("data/output/contract_service_area.rds")
```
```{r, include=FALSE}
final.data1 <- full.ma.data2 %>%
  inner_join(contract.service.area %>% 
               select(contractid, fips, year), 
             by=c("contractid", "fips", "year")) 
```

```{r, include=FALSE}
fig.avg.enrollment<- final.data1 %>% filter(!is.na(avg_enrollment)) %>% group_by(fips, year) %>% 
  select(fips, year, avg_enrollment) %>%
  summarise(all_enroll = sum(avg_enrollment)) %>%
  ggplot(aes(x = year, y = all_enroll)) +
  stat_summary(fun.y = "mean", geom = "bar") +
  labs(x = "Year", y="Average Enrollees", title = "Average Enrollees per County") + 
  scale_y_continuous(labels = comma) + theme_bw()
```
```{r}
fig.avg.enrollment
```

# Question 6
```{r, include=FALSE}
final.data<- readRDS("data/output/final_ma_data.rds")
````

```{r, include=FALSE}
fig.avg.premium<- final.data %>% 
  group_by(year)%>% 
  summarize(prem = mean(!is.na(premium)))%>%
  ggplot( aes(year, prem))+
  geom_line(color = 'red')+
  labs( title = 'Average premium over time', x = 'Year', y = 'Average Premium')+
  theme_minimal()
```
```{r}
fig.avg.premium
```

# Question 7

```{r, include=FALSE}
fig.0premium<- final.data %>% 
  filter(!is.na(premium))%>% 
  group_by(year)%>% 
  summarize(perc_0 = (sum(premium == 0)/n() * 100))%>% 
  ggplot( aes(year, perc_0))+
  geom_line(color = 'blue')+
  labs( title = 'Percentage of $0 Premium Plans over time', x = 'Year', y = 'Percentage of $0 Premium Plans')+
  theme_minimal()
```

```{r}
fig.0premium
```

# Question 8

# Question 9

# Question 10

Working with this data was a little difficiult as I was not familiar with the dataset and some of the terms used. I am also not as familiar with R so there is a steep learning curve for me. It was frustrating when the codes would not run or take a long time to learn as it made it difficult to understand what I did wrong. 