---
title: 'Homework 1 - Research in Health Economics'
author: "Ammarah Ahmed"
date: ""
output: pdf_document 
    
---
```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, readr, 
               scales, kableExtra)
```

# Question 1
```{r, include=FALSE}
full.ma.data<- readRDS("data/output/full_ma_data.rds")
```
```{r, include=FALSE}
total_obs<- as.numeric(count(full.ma.data %>% ungroup()))
```

```{r}
paste0("There are ", total_obs , " observations in the dataset")
```

# Question 2 
```{r, include=FALSE}
plan_type_table<- full.ma.data %>% group_by(plan_type) %>% count() %>% arrange(-n)
```

There are 26 different plans in the dataset.

# Question 3

```{r, include=FALSE}
plan.type.year1<- full.ma.data %>% group_by(plan_type, year) %>% count() %>% 
  arrange(year, -n) %>% filter(plan_type!="NA")
```
```{r, include=FALSE}
pe.year1<- pivot_wider(plan.type.year1, names_from = "year", values_from = "n", names_prefix = "Count_")
```

```{r}
options(knitr.kable.NA = 0)
knitr::kable(pe.year1, 
             col.names=c("Plan Type","2007", "2008", "2009",  "2010","2011","2012","2013","2014","2015"),
             format.args = list(big.mark=","), caption = "Plan Count by Year")
```

# Question 4

```{r, include=FALSE}
full.ma.data2<- full.ma.data %>% 
  filter(snp == 'No' & eghp == 'No' & !(planid %in% 800:899))
```
```{r, include=FALSE}
plan.type.year2<- full.ma.data2 %>% group_by(plan_type, year) %>% count() %>% 
  arrange(year, -n) %>% filter(plan_type!="NA")
```
```{r, include=FALSE}
pe.year2<- pivot_wider(plan.type.year2, names_from = "year", values_from = "n", names_prefix = "Count_")
```

```{r}
options(knitr.kable.NA = 0)
knitr::kable(pe.year2, 
             col.names=c("Plan Type","2007", "2008", "2009",  "2010","2011","2012","2013","2014","2015"),
             format.args = list(big.mark=","), caption = "Plan Count by Year")
```

# Question 5

```{r, include=FALSE}
contract.service.area <- readRDS("data/output/contract_service_area.rds")
```
```{r, include=FALSE}
final.data1 <- full.ma.data2 %>%
  inner_join(contract.service.area %>% 
               select(contractid, fips, year), 
             by=c("contractid", "fips", "year")) 
```

```{r, include=FALSE}
fig.avg.enrollment<- final.data1 %>% filter(!is.na(avg_enrollment)) %>% group_by(fips, year) %>% 
  select(fips, year, avg_enrollment) %>%
  summarise(all_enroll = sum(avg_enrollment)) %>%
  ggplot(aes(x = year, y = all_enroll)) +
  stat_summary(fun.y = "mean", geom = "bar") +
  labs(x = "Year", y="Average Enrollees", title = "Average Enrollees per County") + 
  scale_y_continuous(labels = comma) + theme_bw()
```
```{r}
fig.avg.enrollment
```

# Question 6
```{r, include=FALSE}
final.data<- readRDS("data/output/final_ma_data.rds")
````
```{r, include=FALSE}
fig.avg.premium<- final.data %>% group_by(year) %>% 
  summarise(prem = mean(!is.na(premium))) %>%
  ggplot(aes(x = year, y = prem)) +
  stat_summary(fun.y = "mean", geom = "bar") +
  labs(x = "Year", y="Average Premium", title = "Average Premium over Time") + 
  scale_y_continuous(labels = comma) + theme_bw()
```
```{r}
fig.avg.premium
```

# Question 7

```{r, include=FALSE}
fig.0.premium<- final.data %>% filter(!is.na(premium)) %>% group_by(year) %>% 
  summarise(perc_0 = (sum(premium == 0)/n() * 100)) %>%
  ggplot(aes(x = year, y = perc_0)) +
  stat_summary(fun.y = "mean", geom = "bar") +
  labs(x = "Year", y="Percentage $0 Premium Plans", title = "Percentage of $0 Premium Plans over Time") + 
  scale_y_continuous(labels = comma) + theme_bw()
```

```{r}
fig.0.premium
```

# Question 8

We dropped the "800-series" plans as those represent people who are part of a union and thus do not represent the regular Medicare Advantage user.

# Question 9

Many Medicare Advantage plans offer $0 premiums as the insurance plans may have private contracts with certain health care providers to save costs and may provide preventative care leading to healthier patients who are likely to have lower health costs. While the premium may be $0, the patients may hav other out-of-pocket payments such as copays or deductibles. Also, beneficiaries are likely to be restricted to the health care providers covered by the plan as they will have to pay more to visit doctors outside the plan. 

# Question 10

Working with this data was a little difficiult as I was not familiar with the dataset and some of the terms used. I am also not as familiar with R so there is a steep learning curve for me. It was frustrating when the codes would not run or take a long time to learn as it made it difficult to understand what I did wrong. 